Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Mechanical Perturbations

### 2024-05-31 main.tex
Some todo and image name fix.

\section{Introduction}
%
Single-actor bicycle crashes are associated with a suprisingly large number of
reported injuries~\citep{Wegman2024}. At low speeds, from start up to typical
cruising speeds, bicycles themselves are not self-stable and can be challenging
for the rider to balance. Low speed crashes may be reduced if the bicycle was
self-stable at these speeds by  relieving the rider of a necessary control
activity. Bicycles can be mechanically modified to lower the speeds at which
they are self-stable~\citep{Astrom2005} but since the 1980s, it has been known
that applying a motor acutated steering torque proportional to the roll rate
can stabilize a single track vehicle at very low speeds. If this automatic
control of steering can stabilize a bicycle, it may reduce the control required
by the rider to successfully manage balancing tasks. We have developed such a
balance assisting bicycle and hypothesize that it helps the rider in situations
they are likely to fall.

@@

\section{Methods}
%
\subsection{Bicycle}
%
We modified a standard electric bicycle (Gazelle Grenoble C7, Dieren, The
Netherlands) with a custom motor in the steering assembly capable of applying
up to 7~\si{\newton\meter} of torque between the head tube and steer tube,
Figure~\ref{fig:balance-assist-bicycle}. A custom motor controller converts the
commanded torque to applied torque. We measure the speed of the rear wheel with
an encoder (Magura XXX) and measure the body fixed roll rate of the bicycle
with a MEMs rate gyroscope (Company, Location). The balance assist control
algorithm is implemented on a microprocessor (Teensy, Company, Location) and
data from all sensors is logged with a CAN bus (Company, Location) at
approximately X Hz.
%
\begin{figure}
  \centering
  \includegraphics[width=100mm]{figures/balance-assist-bicycle.jpg}
  \caption{Balance assist bicycle prototype with steer motor and data
    aquisition and control electronics mounted in the rear rack.}
  \label{fig:balance-assist-bicycle}
\end{figure}

@@

\subsection{Measurements}
%
We measure the time histories of the Bump'Em delivered perturbation forces and
the bicycle's steer angle, roll angle, roll angular rate, and forward speed.
Figure~\ref{fig:perturbation} shows an example perturbation force measurement
compared to our Bump'Em controller command.
%
\begin{figure}
  \centering
  \includegraphics[width=140mm]{figures/perturbation_0.png}
  \caption{Motor force and resulting  steer torque perturbation based on an 80N
  counterclockwise applied force.}
  \label{fig:perturbation_0}
\end{figure}

>>>>

\section{Introduction}
%
Single-actor bicycle crashes are associated with a suprisingly large number of
reported injuries~\citep{Wegman2024}. At low speeds, from start up to typical
cruising speeds, bicycles themselves are not self-stable and can be challenging
for the rider to balance. Low speed crashes may be reduced if the bicycle was
self-stable at these speeds by  relieving the rider of a necessary control
activity. Bicycles can be mechanically modified to lower the speeds at which
they are self-stable~\citep{Astrom2005} and since the 1980s it has been known
that applying a motor acutated steering torque proportional to the roll rate
can stabilize a single track vehicle at very low speeds. If this automatic
control of steering can stabilize a bicycle, it may reduce the control required
by the rider to successfully manage balancing tasks. We have developed such a
balance assisting bicycle and hypothesize that it helps the rider in situations
they are likely to fall.

@@

\section{Methods}
%
\subsection{Bicycle}
%
We modified a standard electric bicycle (Gazelle Grenoble C7\todo{Double check
this is the right model}, Dieren, The Netherlands) with a custom motor in the
steering assembly capable of applying up to 7~\si{\newton\meter} of torque
between the head tube and steer tube, Figure~\ref{fig:balance-assist-bicycle}.
A custom motor controller converts the commanded torque to applied torque. We
measure the speed of the rear wheel with an encoder (Magura XXX) and measure
the body fixed roll rate of the bicycle with a MEMs rate gyroscope (Company,
Location). The balance assist control algorithm is implemented on a
microprocessor (Teensy, Company, Location) and data from all sensors is logged
with a CAN bus (Company, Location) at approximately X Hz.
%
\begin{figure}
  \centering
  \includegraphics[width=100mm]{figures/balance-assist-bicycle.jpg}
  \caption{Balance assist bicycle prototype with steer motor and data
    aquisition and control electronics mounted in the rear rack.}
  \label{fig:balance-assist-bicycle}
\end{figure}

@@

\subsection{Measurements}
%
We measure the time histories of the Bump'Em delivered perturbation forces and
the bicycle's steer angle, roll angle, roll angular rate, and forward speed.
Figure~\ref{fig:perturbation_0} shows an example perturbation force measurement
compared to our Bump'Em controller command.
%
\begin{figure}
  \centering
  \includegraphics[width=140mm]{figures/perturbation_0.png}
  \caption{Motor force and resulting  steer torque perturbation based on an 80N
  counterclockwise applied force.}
  \label{fig:perturbation_0}
\end{figure}
